% Use the University of Michigan thesis class.
\documentclass[thesis,openany]{./tex/thesis-umich}


\newcommand*{\ATLASLATEXPATH}{latex/}
\usepackage{\ATLASLATEXPATH atlasphysics}
\usepackage{ANA-HIGG-2020-16-INT1-defs}

%\usepackage{showframe}
\RequirePackage{doi}
\usepackage{hyperref}
\usepackage{pgffor}
\usepackage{longfigure}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{multirow}
\usepackage[export]{adjustbox}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage{makecell}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{xifthen}
\usepackage{rotating}
\usepackage{etoolbox}
\usepackage{tikz}

\usepackage{siunitx}
% High energy physics
\DeclareSIUnit\micron{\micro\metre}
\DeclareSIUnit\mrad{\milli\rad}
\DeclareSIUnit\gauss{G}
\DeclareSIUnit\eVperc{\eV\per\clight}
\DeclareSIUnit\nanobarn{\nano\barn}
\DeclareSIUnit\picobarn{\pico\barn}
\DeclareSIUnit\femtobarn{\femto\barn}
\DeclareSIUnit\attobarn{\atto\barn}
\DeclareSIUnit\zeptobarn{\zepto\barn}
\DeclareSIUnit\yoctobarn{\yocto\barn}
\DeclareSIUnit\nb{\nano\barn}
\DeclareSIUnit\pb{\pico\barn}
\DeclareSIUnit\fb{\femto\barn}
\DeclareSIUnit\ab{\atto\barn}
\DeclareSIUnit\zb{\zepto\barn}
\DeclareSIUnit\yb{\yocto\barn}




% Title of the thesis
\title{Precision Measurements of Higgs Boson Couplings in the Diphoton Decay Channel with Run-2 of the ATLAS Detector}

% Author name
\author{Garrett Merz}

% Department
\department{Physics}

% Year of completion
\year=2021

% ORCID ID 
\orcid{0000-0003-4737-3931}

% UM Unique Name 
\umuniquename{gwmerz}

% Frontispiece
\frontispiece{ “Questing Physicks isn’t like the Quiet and Queer branches. You can’t do it at home in a comfortable chair—you have to be out in the thick of the business, with your tools on your belt and your heart on your sleeve!” - Catherynne Valente, The Girl Who Fell Beneath Fairyland and Led The Revels There \cite{Fairyland} \\ }

% Default style for front pages
\frontpagestyle{1}

% Dedication
%\dedication{}

% Acknowledgments
\acknowledgments[6]{

The acknowledgements section of this dissertation, perhaps surprisingly, is the hardest for me to write. If left to my own devices, I have no doubt that the list of people I'd want to thank would be as long as the rest of this paper, if not longer. Like the globe held up by the Atlas of myth, my world is perched upon the shoulders of giants. Completing a dissertation at any point is no easy task, let alone in the middle of a pandemic- I could not have done this alone.

To all the members of the Fondue group, past and present- thank you for making graduate school a place where I could give my best and know it was enough, and for helping me learn how to keep my often-very-scattered brain on task. Learning how to thrive as part of a big collaboration like ATLAS has been difficult, but Tom, Yanlin, Allison, Rachel, Dan, Hao, Xueyue, Sabrina, and everyone else- if it weren't for the kindness and support of a group like yours, I don't know if I'd have been stubborn enough to stick around long enough to see this thing through.

To all the other math and science mentors I've had over the years- Richard Hughes, Brian Winer, Robert Perry, Richard Chute, Brett Miller, and so many more: thank you for sharing your passion and your time, and for walking me along your part of this road.

To all the friends who have gotten me through: Lucas, Kaeli, Doug, Abby, Gates, Gaurav, Rob, Iris, Johnathon, Jennifer, Rachel, Karley, Owen, Rory, Matt, Kara, Eric, and countless others- thank you. Having a life full of laughter to come back to at the end of a long science time has made the days go by easier.

To my parents, Gregg and Gretchen, thank you for working so hard to foster a lifelong sense of curiosity in me, in a world that doesn't always value it. Without the endless stream of NOVA specials, public lectures, and museum trips, I would not be the person I am today. Thank you for always taking an interest in the ideas I've devoted the last few years of my life to- the endless stream of questions and articles and excitement has made me a much better science communicator, and has served as a constant reminder of why I like this stuff in the first place.

To my brother Grant, thank you for being such a constant companion throughout my science career- the fact that there are two physicists in the family has made this work so much less lonely. The fact that we're so different and yet so similar has shown me that there are many, many different ways to love the universe and the ways it's put together. I've learned a lot about myself during my time in graduate school, and I hope your time is as illuminating as mine has been (though hopefully less exhausting).

To my grandparents, JoAnn and Ray, Ray and Sue, and Bert and Ron- thank you for your investments of time, energy, resources, and love for all these years. I know you're proud of me, but I'm just as proud of you- I hope when I'm old I can love the younger generations of my family as profoundly as you've loved yours. 

To all the Goodharts: thank you for welcoming me into your family, and for being there to offer support, coffee, and company when I've needed it most. You're all wonderful, thoughtful people, and I'm very grateful for your kindness.

To the rest of my family- cousins, aunts, uncles, and the rest: thank you for always being excited about and interested in the things I do and the person I am. It takes a whole village to make a doctor, and you are all a part of the person I am today.

Most of all, of course, thanks to my partner, Abigail Goodhart. Thank you for helping me to nurture a poet's eye as well as a scientist's, for helping me to better know the kind of person I am and want to be, and for making this hard year (and the ones before it) brighter and more full of joy. Every day with you is a gift- even when the world feels like it's falling to pieces, we have fun. I love you.

This material is based upon work supported by the National Science Foundation under Grant No. DGE 1256260. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author and do not necessarily reflect the views of the National Science Foundation.

}
% This command sets the width of the acknowledgments area as a fraction
% of the total width of the text area.
%\acknowledgmentswidth{0.85}

% Preface
%\preface[2]{ Do I need this?}

% Committee
\committee{ %
Professor Thomas Schwarz, Chair \\
Professor Dante Amidei \\ 
Professor Eric Bell \\
Professor James Liu \\
Professor Junjie Zhu \\ }

% Chair must be entered separately for formatting reasons.
\chair{Thomas Schwarz}

% Commands to hide or show lists of figures, tables, etc.
%\hidelistoftables
%\showlistofprograms
\showlistofappendices

% Definition of any abbreviations used.
\abbreviations{ %
    \acro{CERN}{Center for European Nuclear Research}
    \acro{ATLAS}{A Toroidal LHC ApparatuS}
    \acro{CMS}{Compact Muon Solenoid}
    \acro{LHC}{Large Hadron Collider}
    \acro{LEP}{Large Electron-Positron Collider}
    \acro{CP}{Charge-Parity}
    \acro{CPT}{Charge-Parity-Time}
    \acro{BDT}{Boosted Decision Tree}
    \acro{QCD}{Quantum Chromodynamics}
    \acro{STXS}{Simplified Template Cross-Section}
    \acro{CKM}{Cabbibo-Kobayashi-Masakawa}
    \acro{LO}{Leading Order}
    \acro{NLO}{Next-to-Leading Order}
    \acro{NNLO}{Next-to-Next-to-Leading Order}
    \acro{PDF}{Parton Distribution Function}
    \acro{SM}{Standard Model}
    \acro{BSM}{Beyond the Standard Model}
    \acro{HC}{Higgs Characterization}
    \acro{RF}{Radiofrequency}
    \acro{SPS}{Super Proton Synchrotron}
    \acro{LINAC}{Linear Accelerator}
    \acro{PSB}{Proton Synchrotron Booster}
    \acro{LHCb}{Large Hadron Collider Beauty}
    \acro{ALICE}{A Large Ion Collider Experiment}
    \acro{HL-LHC}{High-Luminosity Large Hadron Collider}
    \acro{ECAL}{Electronic Calorimeter}
    \acro{HCAL}{Hadronic Calorimeter}
    \acro{IBL}{Insertable B-Layer}
    \acro{SCT}{Semiconductor Tracker}
    \acro{TRT}{Transition Radiation Tracker}
    \acro{MBTS}{Minimum Bias Trigger Scintillator}
    \acro{EM}{Electromagnetic}
    \acro{LAr}{Liquid Argon}
    \acro{PTP}{p-Terphenyl}
    \acro{POPOP}{1,4-Bis(5-Phenyloxazol-2-yl) Benzene}
    \acro{HEC}{Hadronic Endcap Calorimeter}
    \acro{FCAL}{Forward Cal}
    \acro{MDT}{Monitored Drift Tube}
    \acro{CSC}{Cathode Strip Champer}
    \acro{RPC}{Resistive Plate Chamber}
    \acro{TGC}{Thin Gap Chamber}
    \acro{L1}{Level 1}
    \acro{CTP}{Central Trigger}
    \acro{L2}{Level 2}
    \acro{HLT}{High Level Trigger}
    \acro{GEANT4}{GEometry ANd Tracking 4}
    \acro{ID}{Identification}
    \acro{JVT}{Jet Vertex Tagger}
    \acro{PFlow}{Particle Flow}
    \acro{XGBoost}{eXtreme Gradient Boost}
    \acro{TMVA}{Toolkit for Multivariate Analysis}
    \acro{TI}{Tight and Isolated}
    \acro{NTI}{Non Tight and Isolated}
    \acro{NNLOPS}{Next to Next to Leading Order Parton Showering}
    \acro{EFT}{Effective Field Theory}
    \acro{CB}{Crystal Ball}
    \acro{DSCB}{Double Sided Crystal Ball}
    \acro{SS}{Spurious Signal}
    \acro{GPR}{Gaussian Process Regression}
    \acro{SBBDT}{Signal-Background Boosted Decision Tree}
    \acro{CPBDT}{Charge-Parity Boosted Decision Tree}
    \acro{ROC-AUC}{Receiver Operating Characteristic- Area Under Curve}
    \acro{UEPS}{Underlying Event and Parton Shower}
    \acro{CLs}{Confidence Levels}
    \acro{LUCID}{LUminosity Cherenkov Integrating Detector}
    \acro{KLFitter}{Kinematic Likelihood Fitter}
    \acro{BAT}{Bayesian Analysis Toolkit}
    \acro{GaSBaG}{Gaussian Smoothing for BackGrounds}
    \acro{RBF}{Radial Basis Function}
    \acro{GP}{Gaussian Process}
    \acro{DoF}{Degrees of Freedom}
}

% Some abstract text
\abstract{ %

In the second run of the Large Hadron Collider, proton-proton collisions were recorded with the ATLAS detector at a center-of-mass energy of 13 TeV, almost twice that of the previous run. This dramatic increase in energy has enabled physicists to target and precisely measure rare production modes of the recently-discovered Higgs boson for the first time. 

Due to its special role in explaining the origin of fermion and boson masses, measuring the various interactions of the Higgs is of high priority to the ATLAS collider physics program. The diphoton decay channel of the Higgs ($H \rightarrow \gamma \gamma$) offers one of the best probes of many such interactions due to its relatively clean decay signature and the ATLAS detector's high-quality photon resolution. Two major physics analyses are discussed in this dissertation, both of which target this decay channel. Both use the full Run 2 dataset gathered by the ATLAS detector, collected during the 2015-2018 data-taking period and corresponding to a time-integrated luminosity of 139 $fb^{-1}$.

The first of these analyses is a dedicated measurement of the CP properties of the top-Higgs Yukawa coupling, targeting Higgs production in association with a top quark pair ($ttH$) as well as Higgs production in association with a single top quark ($tWH$ and $tHjb$). Two Boosted Decision Trees are developed, one to separate $ttH+tWH+tHjb$ signal from QCD continuum diphoton background and another to separate CP-even-like signal events from CP-odd-like signal events. 20 categories are constructed using the outputs of these two decision trees, and a likelihood fit is performed across all categories. An upper limit is placed on the $tH$ production cross-section of 11.6 times the Standard Model expectation, and the observed $ttH$ significance is measured to be 5.2 $\sigma$, marking the first observation of the $ttH$ process in a single Higgs decay channel. The fully CP-odd top Yukawa coupling scenario is excluded with a significance of 3.9 $\sigma$, while the CP mixing angle is constrained to be $|\alpha| \geq 43^{\circ}$ at 95\% confidence level.

In the second analysis, a variety of Higgs production modes are characterized using the Simplified Template Cross-Sections (STXS) framework. In total, the cross-section times the diphoton decay branching ratio is measured in 88 categories corresponding to 27 theoretically-motivated STXS kinematic regions. The inclusive Higgs boson production cross-section in the Higgs boson rapidity range $|y_{H}|< 2.5$ times the diphoton decay branching ratio is measured to be $127 \pm 10$ fb. In addition, the $ggF+bbH$ production cross-section is measured to be $104 \pm 11$, the $VBF$ production cross-section is measured to be $10.7^{+2.1}_{-1.9}$, the $WH$ production cross-section is measured to be $6.4^{+1.5}_{-1.4}$, the $ZH$ production cross-section is measured to be $-1.2^{+1.1}_{-1.0}$, and the $ttH+tH$ production cross-section is measured to be $1.2^{+0.4}_{-0.3}$. The compatibility between the measurement and the expected value corresponds to a p-value of 3\%, a 1.9$\sigma$ deviation from the Standard Model. However, when the $WH$ and $ZH$ processes are combined into a single $VH$ process, its cross-section times branching ratio is measured to be $5.9 \pm 1.4$fb, the compatibility between the  measurement and the expected value corresponds to a p-value of 50\%, and no significant deviation from the Standard Model is observed. Similarly, when the STXS bins are considered individually, the p-value of the measurement is 60\%, and no significant deviation from the Standard Model is observed. In addition, an upper limit is placed on the $tH$ production cross-section of 8.2 times the Standard Model expectation, the strictest limit placed on this process to date.}
%\hideabstractpagenumber

%% DOCUMENT AREA

\usepackage{ amssymb }
\usepackage{ amsmath }
\begin{document}

% ----- Introduction ----- %

\chapter{Introduction} \label{chap:intro}
	\input{sections/intro} 

\chapter{An Overview of the Standard Model} \label{chap:theory_chapter}
	\input{sections/theory_chapter} 

\chapter{The ATLAS Detector} \label{chap:detector_chapter}
	\input{sections/detector_chapter} 

\chapter{Experimental Methods} \label{chap:methods_chapter}
	\input{sections/methods_chapter} 

\chapter{Data and Monte Carlo Samples} \label{chap:datamc_chapter}
	\input{sections/datamc_chapter} 

\chapter{Signal Parameterization, Background Parametrization, and Statistical Methods} \label{chap:sigbkgparam}
	\input{sections/sigbkgparam} 

\chapter{Study of the CP Properties of the Top Quark Yukawa Interaction in $t\bar{t}H$ and $tH$ Events with $H \rightarrow \gamma \gamma$: Selection and Categorization} \label{chap:tthcp_chapter}
	\input{sections/tthcp_chapter} 

\chapter{Study of the CP Properties of the Top Quark Yukawa Interaction in $t\bar{t}H$ and $tH$ Events with $H \rightarrow \gamma \gamma$: Results} \label{chap:tthcp_results}
	\input{sections/tthcp_results} 

\chapter{Measurements of the Properties of Higgs Boson Production with $H \rightarrow \gamma \gamma$} \label{chap:couplings_chapter}
	\input{sections/couplings_chapter} 

\chapter{Conclusion} \label{chap:conclusion}
	\input{sections/conclusion} 



% ----- Appendices ----- % 

\appendix

\chapter{Alternative Top Reconstruction with the KLFitter} \label{app:KLFitter}
	\input{appendix/KLFitter}

\chapter{CP-BDT Studies with the Toolkit for Multivariate Analysis} \label{app:TMVABDT}
	\input{appendix/TMVABDTStudies}

\chapter{Auxiliary Plots} \label{app:couplings_auxplots}
	\input{appendix/couplings_auxplots}

\chapter{Reducing Spurious Signal With Gaussian Process Regression} \label{app:gpr_templates}
	\input{appendix/gpr_templates}

\chapter{Validation Tests of the Gaussian Process Regression Method} \label{app:gpr_validation}
	\input{appendix/gpr_validation}


% Using AIAA bibliography style 
\bibliographystyle{aiaa}
% Give this command the relative path to the .bib file.
\bibliography{./bib/thesis-bib}


\end{document}


