\section{The Large Hadron Collider} \label{sec:LHC} 

At roughly 27 kilometers in circumference, the Large Hadron Collider (LHC) at the Center for European Nuclear Research (CERN) is the largest machine ever built, slicing through the solid rock of the Jura mountains and running more than 100 meters beneath the Swiss-French border \ref{LHC}. It is a proton-proton collider operating at a center-of-mass energy of 13 TeV, circulating two beams of protons in opposite directions, each at more than 99\% the speed of light \ref:{speed}. By utilizing a system of radiofrequency (RF) cavities and dipole magnets, protons are delivered to four primary collision points in bunches spaced 25 nanoseconds apart \ref:{40Mhz}.

The LHC achieves such a powerful center-of-mass energy by repurposing a significant fraction of older collider physics infrastructure, including the Super Proton Synchrotron (SPS), which was itself one of the most powerful particle accelerators in the world at one time. In order to achieve collision, hydrogen atoms are first stripped of their electrons using an ionizing cathode filament in a device called a 'Duoplasmatron'; this produces a plasma that is filtered to produce beams of protons \ref{}. The LINAC2, a linear accelerator, then accelerates the resultant proton beams to a collision energy of 50 MeV; the beams then move to the Proton Synchrotron Booster (PSB), a circular accelerator that accelerates them to an energy of 1.4 GeV, followed by the Proton Synchrotron (PS), a second circular accelerator that accelerates them to 25 GeV. Following this, beams enter the aforementioned Super Proton Synchrotron, where they are accelerated to a collision energy of 450 GeV before being injected into the Large Hadron Collider ring. This infrastructure is depicted in Figure \ref{fig:LHC}.

The tight bunching of protons by the LHC ring allows the collider to deliver collisions at a high luminosity, a measure of the number of expected collision events per unit of beam area. Luminosity is measured in both instantaneous and integrated form (i.e., 'luminosity per unit time' and total luminosity delivered to date'); as of the end of the second major LHC run, the collider has delivered a total integrated luminosity of XXX fb^{-1}, equivalent to X.XX*10^{41} collisions per square centimeter \ref{lumi}. 

At the time of this writing, the LHC is nearing the end of 'Long Shutdown 2' (LS2), a two-year long upgrade period designed to increase detector performance and lay the groundwork for the upcoming high-luminosity LHC (HL-LHC) upgrade, which is slated to begin in approximately 2027 and aims to increase the design luminosity of the LHC by at least a factor of 5 \ref:{HL-LHC}. Higher luminosity allows for the production of more rare physics events, but also dramatically increases the incidence of unwanted "pileup" events; mitigating this is one of the major efforts of the CERN physics community during the lead-up to the HL-LHC run.

The LHC ring accelerates beams to a center-of-mass collision energy of 13 TeV. The ring consists of two primary beam pipes, one containing a "clockwise" beam and the other containing a "counterclockwise" beam, which overlap at the four primary collision points. These correspond to the four major LHC physics experiments: ATLAS  \ref{} and CMS \ref{}, which are "general purpose" physics detectors, LHCb \ref{}, which is specialized to study the physics of hadrons containing bottom (or "beauty") quarks, and ALICE \ref{}, which is primarily designed for heavy-ion physics. Due to the highly sensitive nature of their instrumentation, these detectors are buried approximately 100 meters underground to avoid interference from high-energy cosmic rays.

\section{The ATLAS Detector} \label{sec:ATLAS} 

The work contained in this dissertation was performed using the ATLAS (A Large Toroidal LHC ApparatuS) detector. Together with its "sibling detector" CMS (the Compact Muon Solenoid), the ATLAS detector is one of CERN's two "general purpose" detectors, so-called because it is designed to observe a wide variety of high-energy physics phenomena. The ATLAS detector is cylindrical, with a diameter of 25 meters and a length of 44 meters \ref{ATLAS}. Upon colliding at the interaction point in the center of the detector, each proton's constituent particles will interact with one another through a host of different physical processes, producing a variety of new particles traveling at high velocities. The ATLAS detector is comprised of several specialized subsystems, each of which is responsible for logging different properties of these decay products. Each of these subsystems is optimized to detect specific particles and properties; taken together, they provide a detailed picture of a collision event much like different color channels in a photographic camera.
	The Inner Detector (ID), a high-resolution detector composed of silicon "strips", "pixels", and "straws", measures the tracks of particles close to the primary vertex. It is encased within a 2T solenoidal magnetic field that bends the tracks of charged particles produced at the collision point, allowing their momenta to be measured as a function of the curvature of their trajectories. Enclosing the Inner Detector are two nested calorimeters: one scintillator-tile electronic calorimeter (ECAL) that measures the energy of incident photons and electrons, and one liquid-argon hadronic calorimeter (HCAL) that measures the energy of hadronic showers. Finally, a muon spectrometer composed of gas-filled tubes and chambers records the tracks of muons, which do not decay in the calorimeters like most other particles due to their status as minimum-ionizing particles \ref{muons}. The muon spectrometer has three major subsystems, a cylindrical barrel and two circular endcaps, which provide spatial coverage in both the central and forward regions of the ATLAS detector. The entirety of the muon spectrometer is suffused with magnetic field generated by 24 toroidal magnets, eight of which lie in the barrel and generating a field of 1T and eight of which lie in each endcap generating a field of 0.5T. These magnets bend the trajectories of muons in a manner similar to the inner solenoidal magnet system, allowing the momenta of particles to be measured based on the track curvature. The ATLAS detector is also equipped with a highly sensitive trigger system, which uses a wide variety of physics signatures in each of these detector subsystems to determine which collision events to store for later analysis \ref{ATLAS}.
	The ATLAS detector uses a right-handed coordinate system, with the central interaction point at z = 0. The x-axis points from the interaction point to the center of the LHC ring, while the y-axis points upward. However, particle trajectories are more commonly measured not in (x,y,z) coordinate space but utilizing angular variables (\(r\) and \( \phi \)) in the transverse plane and pseudorapidity \( \eta \) in the longitudinal plane, where \( \eta = -ln(tan( \theta /2) ) \). This is illustrated in figure \ref{fig:coords}.

\subsection{Inner Detector} \label{sec:ID} 

The Inner Detector is composed of five highly granular subdetectors. The closest to the beam, the silicon-based Insertable B-Layer, was added between the LHC's first and second runs in order to improve precision track-vertex measurement for tasks such as b-hadron tagging and \( \tau\) lepton identification \ref{IBL}, methods discussed more in section \ref{tagging}. Proceeding outward from the IBL are the silicon-based Pixel detector and Semiconductor Tracker (SCT), followed by the gas-based Transition Radiation Tracker (TRT) \ref{ATLAS}. 

The IBL is composed of 26,880 250×50 \(μm^2\) silicon pixels arranged in an 80 column by 336 row geometry. The pixel modules are supported by 14 support staves, tilted at approximately 14 degrees from the nominal to achieve near-complete cylindrical coverage of the barrel region. The IBL is approximately 33.25 mm from the beam axis, necessitating a replacement of the beryllium beam-pipe with a smaller-radius one for detector reintegration \ref{IBL}.

Silicon is an optimal material for precision trackers due to its status as a semiconductor. Silicon wafers can be easily "doped" with atoms of other elements, leading them to carry either an excess of electrons (n-type semiconductors) or a deficit of electrons (p-type semiconductors). By putting the two together, an n-p junction is formed, leading to a one-way current gate known as a diode. A voltage is then applied to the diode to halt current flow ("reverse biasing"). When ionizing particles such as those produced in collisions pass through the silicon, they will produce a cascade of charge carriers that will induce a current in the diode, which can then be measured by the precision front-end electronics attached to the silicon wafers \ref{Knoll Radiation Detection}.

The pixel detector is similarly composed of silicon pixels, each of which is either 50x400 \(μm^2\) (nominal) or 50x600 \(μm^2\) (near the front-end readout electronics due to spatial constraints). The detector contains 1744 identical pixel sensors, each of which contains 47232 pixels. Three layers of pixels in the barrel region sit at r = 50.5,  88.5 and 122.5 mm from the beam axis, while three layers of pixels in the endcap region sit at r= 495, 580, and 650 mm from the beam axis. The silicon wafers in the pixel detector are "n+ type" strips (n-type semiconductors with extra doping) embedded in an n-type bulk, rather than the traditional n-type and p-type semiconductors due to the fact that excessive radiation from the beam can cause standard n-and-p-type semiconductors to change into one another. 

The SCT is composed of longer silicon microstrip wafers made out of standard n-p type semiconductors. It consists of four layers of 2112 total silicon microstrip modules in the barrel region and 1976 microstrip modules in the endcap regions, with 770 microstrips per sensor module. The strips are 12cm long and 80 μm wide; the distance of modules from the beam pipe ranges from radii of r= 299 mm to r= 514 mm in the barrel and r=852.8 mm to r= 2720.2 mm in the endcaps. The SCT is designed using less-costly materials than the TRT and IBL, but serves a valuable complementary function.

Finally, the TRT consists of polyimide "straws" filled with a gas mixture (70\% Xenon, 27\% Carbon Dioxide, and 3\% Carbon Tetrafluoride), each 4mm in diameter and 144 cm long. In the center of each straw tube is a 31 μm in diameter grounded gold-plated tungsten wire filament that serves as an anode. Charged particles travelling through the gas medium will ionize it, producing a cascade of ions that will drift to the anode wire. The TRT is uniquely designed to detect low-energy X-ray transition radiation photons, which allows for better identification of electron candidates \ref{TRT}. The barrel region consists of 73 layers of sensor modules ranging from r= 554 mm to r = 1082 mm from the beam pipe, while the encaps straws, arranged in a "wheel" shape, stretch from r= 852.8 mm to r= 2720.2 mm from the beam pipe.

Together, these subsystems form the ATLAS detector's Inner Detector. Each subdetector has a different spatial resolution- the IBL has a resolution of XXXX, the pixel detector has a resolution of XXXX, the SCT has a resolution of XXXX, and the TRT has a resolution of XXXX. The combined effect of these subdetectors is thus to record particle tracks produced in collisions with very high granularity, allowing for sophisticated physics analyses to be performed. 

\subsection{Calorimeters} \label{sec:Calos} 

This is a subsection. It will be included in the table of contents. 


\subsubsection{ECAL} \label{sec:ECAL} 

\subsubsection{HCAL} \label{sec:HCAL} 




\subsection{Muon Spectrometer} \label{sec:Musyst}
 






\subsection{Magnets} \label{sec:Magnets}


\subsubsection{solenoid} \label{sec:solenoid} 

\subsubsection{toroid} \label{sec:toroid} 




\subsection{Trigger} \label{sec:Trigger}
Due to the high rate of collision events delivered by the accelerator 

Additionally, two Minimum Bias Trigger Scintillators, composed of plastic scintillator tiles, sit on the circular endcaps of the ID in order to provide trigger capabilities in the forward region \(|\eta| > 2.5 \). The MBTS endcap scintillator consists of 2 cm thick polystyrene scintillator disks mounted on each endcap at a distance of 3.6m from the collision point at. 
